### 2024-08-05 paper/paper.tex
first draft

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{array}
\usepackage{multirow}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Analysis of Programming Students' Prompts to Identify Patterns of
Critical Thinking in Interactions with Generative AI \\
{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
should not be used}
\thanks{Identify applicable funding agency here. If none, delete this.}
}

\author{\IEEEauthorblockN{1\textsuperscript{st} Rodrigo Prestes Machado}
\IEEEauthorblockA{\textit{dept. Informática} \\
\textit{Instituto Federal do Rio Grande do Sul}\\
Porto Alegre, Brazil \\
0000-0003-0428-6387}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
}

\maketitle

\begin{abstract}
This document is a model and instructions for \LaTeX.
This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
or Math in Paper Title or Abstract.
\end{abstract}

\begin{IEEEkeywords}
component, formatting, style, styling, insert
\end{IEEEkeywords}

\section{Introduction}

%What is the problem to be solved?

According \cite{Kwan24} further research about the use of AI tools in
education is needed to understand how these tools can be used to improve
students' learning experience. In particular, the study suggested that future
research on students' study habits.

Problems

The literature suggests a concern regarding the impact of ChatGPT on students'
critical thinking abilities. Vargas \cite{Vargas2023} indicated that
ChatGPT use might lead to an overreliance on the tool, potentially diminishing
students' critical thinking skills. Similarly, Cai \cite{cai23} identified
overdependence and reduced intellectual engagement as significant drawbacks of
ChatGPT-supported learning. Chan \cite{chan23} also noted that this
reliance could result in a decrease in critical thinking, as students might make
decisions based solely on the information provided by ChatGPT. Collectively,
these studies highlight a common theme of reduced critical engagement and
overreliance on ChatGPT in educational contexts.

% Are there any existing solutions?

% Which is the best?

% What is its main limitation?

% What do you hope to achieve?

\section{Background}

The paper \cite{Kwan24} examined students from different demographic groups
using AI tools like ChatGPT, comparing perceptions and impacts
among these groups. The results showed that AI tools had a positive impact on
academic performance, especially for students with less prior experience.

\subsection{Critical Thinking}

Critical thinking is a fundamental skill for students to develop, as it enables
them to analyze, evaluate, and synthesize information. This skill is essential
for problem-solving and decision-making, making it a crucial aspect of
educational success, specifically in programming courses. However, the use of AI
tools like ChatGPT has raised concerns about the potential impact on students'
critical thinking abilities \cite{Vargas2023} \cite{cai23} \cite{chan23}.

In the other hand, the study of \cite{zhang24} suggested that incorporating
ChatGPT could stimulate students' critical thinking, probably because they
had to assess the correctness of ChatGPT's answers.

Regarding self-regulation, \cite{cai23} noted that ChatGPT allowed students to
regulate their learning pace.

\cite{wu24} found that ChatGPT-supported learning not only promoted students'
self-regulation but also increased their self-efficacy.

Finally, some students asserted that the use of ChatGPT enhanced the depth of
their thinking and understanding \cite{chan23}.

\subsection{CharlieBot}

CharlieBot is a ChatGPT-based bot that provides assistance to students in Java
programming courses at UC3M. The bot uses RAG(retrieval augmented generation) to
generate responses to student queries, providing explanations, code snippets,
and debugging help. The goal of CharlieBot is to support students in their
learning process by providing personalized assistance and feedback.

\begin{table*}[htbp]
\caption{Categories of Analysis}
    \begin{center}
    \begin{tabular}{|p{3cm}|p{5cm}|p{3cm}|}
    \hline
    \textbf{Category} & \textbf{Description} & \textbf{Critical Thinking} \\
    \hline
    Debugging Help & Prompts that seek help to identify, fix errors, or understand the provided code snippet. & \multirow{6}{3cm}{Tends to promote.} \\
    Conceptual Questions & Prompts that are more about understanding concepts or algorithms rather than specific code. & \\
    Student Correction & Prompts where the student corrects the bot. & \\
    \hline
    Code Snippet & Prompts that ask for a specific part of the code, like a function or a segment. & \multirow{6}{3cm}{Tends not to promote.} \\
    Complete Solution & Prompts that request an entire solution or a complete code snippet. & \\
    Multiple Question & Prompts where the user wants to solve a multiple-question exercise. & \\
    \hline
    Language Change & Prompts that request a change of idiom to convey a message more effectively. & \multirow{3}{3cm}{Neutral.} \\
    Uncategorized & Prompts that do not fit into any of the above categories. & \\
    \hline
    \end{tabular}
    \label{tab:categories}
    \end{center}
\end{table*}

The table \ref{tab:categories} shows the categories of analysis used in this
study. The categories were based on the work of
\cite{Ghimire24} with the addition of four new categories.
Also, we group the categories to understand the impact on critical thinking.

\section{Method}

CharlieBot is a ChatGPT-based bot that provides assistance to students in
Java programming courses in UC3M. CharlieBot logs all conversations with
students, which were used for this study. The goal was to categorize student
messages to understand how students interact with the bot and identify patterns
in their study habits.
\\
The categories outlined in \cite{Ghimire24} served as the
foundation for classifying student messages (Debugging Help, Conceptual
Questions, Code Snippet and Complete Solution). However, it was identified that
some categories were missing, necessitating the addition of new categories
(Student Correction, Multiple Question, Language Change and Uncategorized). Table
\ref{tab:categories} presents these categories along with their descriptions.
\\
Subsequently, these eight categories were input into Claude.ai, accompanied by
detailed instructions on the desired structure of the input and the expected
format of the output. For each conversation with the CharlieBot log, a .csv file
was downloaded, and Claude was tasked with classifying the user messages into
one of the eight categories.

The classifications made by Claude were then manually reviewed. In cases where
there was disagreement with the classification, manual adjustments were made,
and the accuracy of Claude's classification was recorded. Each conversation
with CharlieBot was saved into a .csv file containing the original conversation,
along with two additional columns: "Classification" to denote the category and
"AI" to indicate Claude's accuracy.

As new categorizations were completed, an example was submitted to Claude's
knowledge base to enhance accuracy. Initially, Claude began with an example
representing 1\% of the total memory of its knowledge base. As classifications
were progressively completed, more examples were provided to Claude to increase
the percentage of its knowledge base occupied by relevant examples.

Upon completion of all categorizations, all resulting .csv files were consolidated
for final counts and analysis using the Pandas library. This methodology
facilitated an efficient categorization of student messages, thereby enabling
a comprehensive analysis of how tools such as ChatGPT are utilized in foundational
programming courses.

\section{Results and Discussion}

The graph \ref{fig:graph1} shows the distribution of the messages in the
categories:
\begin{figure}[h!]
    \centering
    \includegraphics[scale=0.7]{figures/figure1.png}
    \caption{Classification of messages.}
    \label{fig:graph1}
\end{figure}

Approximately 64\% of the messages pertain to categories associated with
critical thinking, corroborating with the findings of
\cite{Ghimire24}. In contrast, around 28\% of the messages
indicate a preference among students for ready-made answers.

\section{Conclusion}

\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}