### 2025-01-21 CSEDU2025/Paper.tex
paper update

\documentclass[a4paper,twoside]{article}

\usepackage{epsfig}
\usepackage{subcaption}
\usepackage{calc}
\usepackage{amssymb}
\usepackage{amstext}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{multicol}
\usepackage{pslatex}
\usepackage{algorithm2e}
\usepackage[bottom]{footmisc}
% Please add other packages that you may need BEFORE the SCITEPRESS.sty package.
\usepackage{natbib}
\usepackage{multirow}
\usepackage{SCITEPRESS}
\usepackage{arydshln}
\widowpenalty=10000
\clubpenalty=10000

\begin{document}

\title{
  Analyzing Student Use of Spacing and Interleaving Strategies in Interactions
  with GenAI-Powered Chatbots in Programming Courses
}

\author{\authorname{Rodrigo Prestes Machado\sup{1}\orcidAuthor{0000-0003-0428-6387},
Carlos Alario-Hoyos\sup{2}\orcidAuthor{0000-0002-3082-0814},
Patricia Callejo\sup{2}\orcidAuthor{0000-0001-6124-6213},
Iria Estévez-Ayres\sup{2}\orcidAuthor{0000-0002-1047-5398},
Carlos Delgado Kloos\sup{2}\orcidAuthor{0000-0003-4093-3705}}
\affiliation{\sup{1}Department of Informatics, Instituto Federal de Educação,
Ciência e Tecnologia, Porto Alegre, Brazil}
\affiliation{\sup{2}Department of Telematics Engineering, Universidad Carlos III
de Madrid, Leganés (Madrid), Spain}
\email{\ rodrigo.prestes@poa.ifrs.edu.br, \{calario, pcallejo, ayres, cdk\}@it.uc3m.es}
}

\keywords{
  Programming, Generative Artificial Intelligence, Chatbots, Spacing,
  Interleaving and Metacognition
}

\abstract{
This study aimed to analyze prompts of programming students with a chatbot
powered by OpenAI's GPT-3.5, enhanced with the Retrieval Augmented Generation
(RAG) technique, within the context of a Java programming course. The focus was
on students using two metacognitive strategies: interleaving and spacing.
Student prompts were categorized into eight categories along with their
respective study topics. Findings revealed that the categories and markers of
spacing and interleaving were important in identifying study sessions with the
chatbot. However, students showed limited intentional application of these
learning strategies. These results highlight the need for more comprehensive
guidance on leveraging AI tools to improve learning outcomes.
}

\twocolumn\maketitle\normalsize\setcounter{footnote}{0}

\section{\uppercase{Introduction}}
\label{sec:introduction}

Recent advances in Generative Artificial Intelligence (GenAI) have created
new opportunities in both professional programming \citep{Peng23}
\citep{Pandey24} and education \citep{Puryear22}. In programming courses,
students can employ GenAI tools to improve their understanding, receive
personalized feedback, and access detailed explanations. For example, GitHub
Copilot, a tool that is integrated into development environments, assists in
providing real-time suggestions and accelerating code writing, which could help
streamline the learning process \citep{Denny23}. Educational chatbots, on the
other hand, try to guide students through coding challenges, answer questions,
and try to foster a deeper understanding of programming concepts
\citep{Labadze23}.

The study conducted by \cite{chan23} revealed that both undergraduate and
postgraduate students have positive attitudes towards the use of GenAI.
A systematic review organized by \cite{Lo24} demonstrated that students can
effectively learn from ChatGPT, resulting in improved comprehension and academic
achievement \citep{Callejo24}. Furthermore, it was observed that ChatGPT may
allow students to regulate their learning pace \citep{Baha24} and can
potentially support self-regulated learning, particularly for students with
prior technical and disciplinary knowledge \citep{Xia23}.

However, researchers have also expressed concerns about the impact of these
tools on students. The systematic review by \cite{Murillo23} indicated that the
use of ChatGPT could lead to overreliance on the tool. \cite{chan23} noted that
overdependence on ChatGPT could lead to a decrease in critical thinking, as
students can make decisions based solely on the information provided by the
tool. As a potential consequence, \cite{Bastani24} observed in a study that when
programming students lost access to ChatGPT, those who had previously relied
on it saw their performance drop by 17\%. In contrast, students who had never
used the tool were unaffected and outperformed their peers.

The confidence of students in these tools is well founded, as shown by
\cite{Puryear22}, who found that GitHub Copilot can generate solutions for
student assignments with precision rates ranging from 68\% to 95\%. However,
\cite{Boudouaia24} observed that using ChatGPT without a structured learning
approach did not provide significant improvement over traditional self-directed
learning methods in programming. This raises concerns that students may become
overly dependent on generative AI tools, potentially hindering their learning
progress and missing opportunities to develop a deeper understanding of
fundamental concepts.

Regardless of teachers' preferences or beliefs, preliminary surveys conducted
by \cite{Dickey24} indicate that more than 54.5\% of students are already using
GenAI for homework, likely due to their perception of this technology as
advantageous and intuitive \citep{Boudouaia24}. All of these studies highlight
the need to increase the understanding of how students interact with these tools
and how they can be used to improve learning, as emphasized by \cite{Lo24}.

In response to the growing need for deeper insight into the use of GenAI tools
in educational settings, this study aims to help address this gap by analyzing
the interactions between students in a Java programming course and an
educational chatbot powered by OpenAI gpt-3.5, enhanced with the
Retrieval-Augmented Generation (RAG) technique. Specifically, it focuses on
examining these interactions within the context of two metacognitive strategies:
spacing \citep{Carvalho20} and interleaving \citep{Rivers21}. To achieve this
objective, we formulate three research questions:

\begin{itemize}
  \item RQ1 - What distribution patterns emerged in the classified student
  interaction prompts with the educational chatbot?
  \item RQ2 - Was there spacing between student prompts? Which category led to
  the most rapid and consistent interactions with the chatbot?
  \item RQ3 - Do students' interactions with the chatbot alternate between study
  topics to suggest interleaving?
\end{itemize}

This paper is organized as follows. Section 2 presents the theoretical
framework, Section 3 describes the material and methods used in the study,
Section 4 presents the results and discussion, and Section 5 concludes the
paper and outlines future work.

\section{\uppercase{Theoretical Framework}}

This section outlines the theoretical framework that is the basis of the study.
We begin by introducing the concept of metacognition, along with the spacing and
interweaving strategies applied in this research. Next, we present related works
organized into two subsections: research involving GenAI and studies focusing
on metacognition.

\subsection{Metacognition}

Given that GenAI often produces highly accurate and automatic responses
\citep{Puryear22}, it is essential that students use these tools within an
active learning process to enhance their learning outcomes. This active
learning process can be further understood through the lens of metacognition,
which focuses on awareness of one's mental processes. \cite{flavell79}
proposed a model of metacognitive monitoring that includes four interrelated
phenomena: Metacognitive Knowledge, Metacognitive Experience, Metacognitive
Goals, and Metacognitive Actions. These processes do not occur in isolation, but
they influence each other, altering cognitive progress over time.

Metacognitive knowledge includes beliefs about variables that affect the
outcomes of cognitive activities. It is divided into three types: beliefs about
personal abilities, perceived difficulty in the task, and previously used
strategies. Metacognitive Experience refers to the feelings that arise before,
during, and after cognitive activity, such as frustration, confusion,
satisfaction, and others. Metacognitive Goals are the key to regulating thought
as they relate to the goals the individual seeks to achieve, directly
influencing the actions taken. For example, if a student’s goal is to complete
a task quickly, they may adopt a more passive approach to learning. Lastly,
Metacognitive Actions involve the planning, monitoring, and evaluation of
strategies used to achieve the goals. In terms of planning, students can
determine how to approach a task, such as spacing their study sessions
\citep{Ouhao18, Carvalho20}, interleaving topics \citep{Rivers21}, utilizing
retrieval practice \citep{larsen18}, and other possible strategies.

Spacing and interleaving are two metacognitive strategies that have been shown
to enhance learning outcomes and support the research questions of this study.
Spacing refers to the practice of distributing study sessions over time, which
has been shown to improve long-term retention and understanding of the material
\citep{Carvalho20}. Interleaving involves mixing different topics or problems
within a single learning session, which has been shown to also enhance long-term
learning and the application of student's knowledge in other contexts and
situations \citep{rohrer20} \citep{Rivers21} .

\subsection{GenAI Related Works}

\cite{Margulieux24} conducted a study on how undergraduate students in
introductory programming courses used generative AI to solve programming
problems in a naturalistic setting. The research focused on examining the
relationship between AI usage and students’ self-regulation strategies,
self-efficacy, and fear of failure in programming. Furthermore, the study
explored how these variables interacted with the characteristics of the learners,
the perceived usefulness of AI, and academic performance. The findings revealed
that students with higher self-efficacy, lower fear of failure, or higher prior
grades tended to use AI less frequently or later in the problem-solving process
and perceived it as less useful compared to their peers. However, no significant
relationship was found between students’ self-regulation strategies and their
use of AI.

% \cite{Boudouaia24} investigated the impact of ChatGPT-facilitated programming on
% college students’ programming behaviors, performance, and perceptions. The
% problem addressed was whether the use of ChatGPT in programming education could
% influence how students interacted with programming tasks, their learning
% outcomes, and their attitudes towards technology. The study employed a
% quasi-experimental design involving 82 college students divided into two groups:
% one using ChatGPT-assisted Programming (CAP) and the other using traditional
% self-directed Programming (SDP). The method combined mixed data collection
% techniques, including behavioral logs, programming performance evaluations, and
% interviews. The results showed that students using the CAP mode exhibited more
% frequent behaviors such as debugging and reading feedback. Although CAP students
% had a higher average score in programming performance, there was no
% statistically significant difference compared to the SDP group. However,
% students’ perceptions of ChatGPT improved significantly after the intervention,
% with increased perceived usefulness, ease of use, and intention to continue
% using the tool in the future.

The study of \cite{Boudouaia24} examined the effects of ChatGPT-assisted
programming on university students' behaviors, performance, and perceptions.
A quasi-experimental research was conducted with 82 students divided into two
groups: one with ChatGPT-assisted programming (CAP) and the other with
self-directed programming (SDP). The analysis included behavioral logs,
performance evaluations, and interviews. Students in the CAP group engaged more
actively in debugging and feedback review activities. Although they achieved
slightly higher scores, there was no statistically significant difference in
performance compared to the SDP group. Nevertheless, perceptions of ChatGPT
improved significantly, highlighting greater perceived usefulness, ease of use,
and intention to use the tool in the future.

\cite{Bastani24} investigated the impact of GenAI, specifically GPT-4, on human
learning, with a focus on mathematics education in a high school. The problem
addressed was how the use of generative AI could affect the acquisition of new
skills, which is crucial for long-term productivity. The method involved a
controlled experiment with approximately one thousand students, who were exposed
to two GPT-4-based tutors: a simple tutor (GPT Base) and another with safeguards
designed to promote learning (GPT Tutor). The results showed that while access
to GPT-4 improved performance on practice exercises (48\% with GPT Base and
127\% with GPT Tutor), the removal of access to GPT Base led to a 17\% decrease
in student performance on exams. This suggests that unrestricted use of
\textit{GPT Base} could hinder learning. However, the GPT Tutor was able to
mitigate this negative effect.

\begin{table*}[htbp]
  \caption{Categories - adapted from \cite{Ghimire24}}
  \begin{center}
    \renewcommand{\arraystretch}{1.1} % Increase the spacing between rows
    \begin{tabular}{p{3.8cm} p{6cm} p{4.5cm}} % Remove vertical bars
      \hline
      \textbf{Category} & \textbf{Description} & \textbf{Example of prompt} \\
      \hline
      Debugging Help (DH) & Prompts that seek help to identify, fix errors, or understand the provided code snippet. & \textit{Would this code be ok? \{code\}} \\
      Conceptual Question (CQ) & Prompts that are more about understanding concepts than specific code. & \textit{What does it mean for a method to be static?} \\
      Student Correction (SC) & Prompts where the student corrects the chatbot. & \textit{The correct answer is B} \\
      Code Snippet  (CS) & Prompts that ask for a specific part of the code, like a function or a segment. & \textit{A class inherits from another write this code} \\
      Complete Solution (CSO) & Prompts that request an entire solution or a complete code snippet. & \textit{Give me the code for a selection sort} \\
      Multiple Question (MQ) & Prompts where the user wants to solve a multiple choice exercise (Quiz). & \textit{A heap is a data structure appropriate for: \{options\}} \\
      Language Change (LC) & Prompts that request a change of idiom. & \textit{In Spanish} \\
      Uncategorized (U) & Prompts that do not fit into any of the above categories. & \textit{Thanks} \\
      \hline
    \end{tabular}
    \label{tab:categories}
  \end{center}
\end{table*}

\subsection{Metacognition Related Works}

Learning based on metacognitive strategies has been shown to be effective in
enhancing students’ performance and skills. \cite{Zheng19} investigated the
effects of metacognitive scaffolds on group behavior, performance, and cognitive
load in computer-supported collaborative environments. The results showed
positive impacts on metacognitive behavior and group performance without
increasing the cognitive load. \cite{LiWei23} proposed a metacognition-based
collaborative approach to enhance performance in collaborative programming. The
findings indicated that this approach influenced computational thinking,
critical thinking, and metacognitive awareness. Similarly, \cite{Wang23}
explored how metacognitive instruction affects computational thinking, critical
thinking, and metacognitive skills in collaborative programming, observing

% \cite{Khusnul24} investigates
% the efficacy of a meta-learning approach in improving metacognitive and creative
% skills. This study confirms the effectiveness of metacognition strategies and
% elucidates the relationship between meta-learning and metacognition. This
% research suggests that meta-learning can improve metacognitive abilities,
% providing valuable insights into educational technology and course design in
% higher education settings.

% \cite{Sidra24} addressed the growing importance of metacognitive skills in an
% AI-enabled workforce. The problem discussed was how AI systems, especially in
% collaborative tools based on Large Language Models (LLMs), impacted human
% decision-making and the challenges they posed due to the lack of human-like
% metacognition in AI. The method used identified four main characteristics that
% differentiated human-AI interaction from human-human collaboration and proposed,
% based on the dual-process theory of human thought, that metacognitive thinking
% was crucial for the effective use of AI. The results indicated that human
% workers needed to develop and apply metacognitive skills, such as planning,
% monitoring, and evaluating their thought processes, to mitigate cognitive biases
% and improve decision-making when working with AI. The study concluded by
% recommending improvements in AI design and training to improve workers'
% metacognitive abilities, ensuring more effective collaboration with AI systems.

\cite{Sidra24} explored the importance of metacognitive skills in an AI-enabled
workforce, focusing on how AI, particularly tools using Large Language Models,
affects human decision-making. They highlighted challenges arising from AI's
lack of human-like metacognition and identified key differences between human
AI and human-human collaboration. Using dual-process theory, they emphasized the
need for workers to develop metacognitive skills like planning, monitoring, and
evaluating to counter cognitive biases and enhance decision-making. The study
recommended improving AI design and training to boost workers' metacognitive
abilities for more effective collaboration.

\section{\uppercase{Material and Method}}

This section outlines the tools and procedures used in this study. CharlieBot
served as the GenAI tool for educational purposes, and the method was structured
into three distinct phases, each designed to systematically assess its
interaction with students.

\subsection{CharlieBot}

CharlieBot is an educational chatbot powered by ChatGPT 3.5 and enhanced with
Retrieval-Augmented Generation (RAG) \citep{Sun24}. RAG is an AI technique that
integrates information retrieval with generative models. It first retrieves
relevant documents or data from a knowledge base or external source using a
retrieval model. Then, a generative model uses this retrieved information to
generate more accurate, context-aware responses. This dual-step process enhances
the quality and reliability of the chatbot's answers. A prior study on
CharlieBot's performance revealed that most students found its responses
appropriate for the Java programming course \citep{Hoyos24}.

\subsection{Method}

The study comprised three phases: data collection, categorization, and analysis.
During the data collection phase, students enrolled in a second-semester Java
course at the University Carlos III of Madrid (UC3M) were introduced to
CharlieBot and allowed to use it without following any prescribed educational
methodology. All data were collected anonymously to ensure that interactions
could not be traced back to individual students or linked to their academic
performance.

During the categorization phase, the students' prompts were initially classified
into eight distinct categories using the Claude.ai \citep{claude} tool.
Subsequently, the authors of this study manually reviewed these classifications
to ensure accuracy and alignment with the research objectives.

Initially, \cite{Ghimire24} proposed four categories: Debugging Help (DH),
Conceptual Question (CQ), Code Snippet (CS), and Complete Solution (CSO).
However, the data collected indicated the need for additional categories,
leading to the inclusion of four more: Multiple Questions (MQ), Student
Corrections (SC), Language Change (LC), and Uncategorized (U). Unlike study
\cite{Ghimire24}, which analyzed students' behavior using artificial
intelligence for specific programming tasks, our study allowed participants to
explore the chatbot as they preferred. This approach enabled the emergence of
additional categories of analysis, broadening the understanding of the tool's
uses and interactions. Table \ref{tab:categories} presents these categories,
their descriptions, and an example. Besides that, the topics covered in the
students' prompts were categorized based on the course issues, which include
(1) Java, (2) Object Orientation, (3) Testing, (4) Recursion, (5) Data
Structures, and (6) Sorting and Searching Algorithms.

The analysis phase involved using Python/Pandas scripts to extract information
from previously classified data. The data sets, and code used in this process
are available at https://github.com/rodrigoprestesmachado/uc3m.

\section{\uppercase{Results and Discussion}}

A total of 625 student messages were categorized in 81 conversations, with
an average of 7.72 messages per conversation. This data offers insights into how
users engage with CharlieBot and the types of prompts they submit. The results
of the analysis are presented in this section, addressing the research questions
outlined in the introduction. Figure \ref{fig:graph1} shows the distribution of
the messages in the categories. The following sections present the results and
discussions of each research question.

\subsection{Categorization of messages}

About the first research question: \textit{RQ1 - What distribution patterns
emerged in the classified student interaction prompts with the educational
chatbot?}

According to Figure \ref{fig:graph1} messages classified as
Conceptual Question account for 35.8\% of the total. Sequences of Conceptual
Questions often start from a practical perspective of a given code, as
illustrated in conversation 4 of Figure \ref{fig:graph2}, which presents
examples of students' interactions with the chatbot. In other cases, the
conversation consists entirely of messages classified as Conceptual Question,
such as in conversation 15 of Figure \ref{fig:graph2}.

\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.61]{img/figure1.png}
  \caption{Categorization of messages.}
  \label{fig:graph1}
\end{figure}

\begin{figure*}[htbp]
  \centering
  \includegraphics[scale=0.41]{img/figure2.png}
  \caption{Examples of conversations.}
  \label{fig:graph2}
\end{figure*}

As shown in Figure \ref{fig:graph1}, approximately 29.9\% of the messages were
classified as Debugging Help. Debugging help messages are likely to provide
students with a practical approach to understanding code, identifying errors,
and building skills to enhance their debugging abilities for future tasks.

Most of the messages are located in the categories Conceptual Question and
Debugging Help corroborate the findings of \cite{Ghimire24} which showed that
the questions are also localized in the same categories.

Multiple Question exercise resolutions represented 14.9\% of the responses.
These prompts typically involved students submitting one or more exercises to
CharlieBot and requesting solutions. The conversation 13 in Figure
\ref{fig:graph2} illustrates a sequence of multiple choice questions.

Requests for the chatbot to generate a Code Snippet or Complete Solution accounted
for 8.2\% of student messages. These prompts reflected a desire for more
direct answers; however, it was observed that, after receiving these solutions,
many students transitioned to a more active approach in seeking to understand
the code or the underlying concepts. Thus, it can be inferred that the prompts
within the Code Snippet or Complete Solution category were often used as a
starting point for a more in-depth study.

Student corrections to chatbot responses represented 4.2\% of the interactions.
In most cases, these corrections occurred when the student already had the
correct answer to an exercise and identified an error in the chatbot response,
as illustrated in conversation 2 in Figure \ref{fig:graph2}. These corrections
highlight that, although the chatbot provides quick feedback, it is not always
accurate.

Figure \ref{fig:graph1} also indicates that approximately 5.8\% of the prompts
written by the students were classified as Uncategorized. These messages include
expressions of gratitude toward the chatbot, such as a simple '\textit{thanks}',
contextual statements like '\textit{I'm reviewing object orientation}', as well
as requests for additional exercises and summaries. These Uncategorized messages
can be subdivided into new categories depending on the investigation. For
instance, requests for new exercises or summaries could be interpreted as a
metacognitive monitoring strategy.

Finally, 1.3\% of the messages were classified as Language Change. These prompts
were typically requests for the chatbot to switch languages (English to Spanish),
as seen in Figure \ref{fig:graph2}, conversation 9, message 2.

\subsection{Spacing}

About the first part of the second research question: \textit{RQ2 - Was there
spacing between student prompts?}

For the spacing analysis, we initially considered a minimum interval of 60
minutes as a reference to define the significant spacing \citep{Gadella24}.
Based on this, our results showed that, of the 81 conversations analyzed, just
27, or around 34\%, had at least one message with an interval more significant
than 60 minutes. Figure \ref{fig:graph2} illustrates the elapsed time between
messages (delta). For example, in conversation 15, the first message was sent
at time zero, while the student sent the second message after a delta of 0.6
minutes. Additionally, message 4 in conversation 15, as shown in Figure
\ref{fig:graph2}, presents a delta of 69.1 minutes, which, according to our
reference, characterizes spacing. The average number of subsections per
conversation was 2.2 among conversations that exhibited spacing. Therefore, a
conversation with one spacing is divided into two distinct study sessions.

This data suggests that there is indeed little spacing between study sessions.
To create a new section in CharlieBot; students must close the browser tab and
start a new conversation, which results in losing the previous chat history.

\subsubsection{Interaction Time per Category}

About the second part of the second research question: \textit{RQ2 - Which
category led to the most rapid and consistent interactions with the chatbot?}

The objective of this research question was to study the time a student takes to
re-engage with the bot after sending a message for each category. This analysis
can offer insights into students' behavior: faster interactions might with the
chatbot might suggest a more passive learning posture.

To calculate the average time and standard deviation for messages within a
category initially was first observed conversations in which message timing was
influenced by previous messages. Pearson's correlation analysis revealed that
28.40\% of the conversations exhibited a strong correlation, which, in this
study, was defined as values outside the range of -0.3 to 0.3. In this context,
a strong correlation suggests that the response time for a message may depend
on the time of the preceding message or multiple prior messages. Consequently,
these 28.40\% of conversations were excluded from the average time and standard
deviation calculation, leaving 56 out of the initial 86 conversations. Table
\ref{tab:averages} presents the means and standard deviations for each category.

When a student requests a Code Snippet (CS), it is generally predictable that
they will summarize interaction with the chatbot within a short period, as
indicated by the mean (1.34) and standard deviation (1.95) of the Code Snippet
category in Table \ref{tab:averages}. This result is noteworthy because,
although some requests are simple and lead to quick responses, others may be
more complex and require more reflection time. However, the low average response
time and minimal variability may indicate that students adopt a more passive
learning posture when requesting the chatbot to generate pieces of code.

Despite occasional variation (3.59), prompts classified as Uncategorized exhibit
a relatively low average time (1.70) for students to resume interaction with the
chatbot. This variation occurs partly because of situations where students end
the conversation with an Uncategorized message like \textit{"thanks"} but later
decide to resume the interaction.

When a student requests the chatbot to Change the Language (LC)  in a response,
in our context English to Spanish, the new interaction occurs quickly (2.23)
with slight variation in response time (3.71), meaning that it is predictable
that the student will continue interacting.

Although still below the overall average, students take moderate time (2.47) to
resume interaction with the chatbot after requesting the solution to a Multiple
Question (MQ) exercise. The variation (5.56) suggests that some responses may
lead to a slower follow-up interaction, possibly indicating a moment of deeper
reflection on the part of the students. However, since the mean and standard
deviation are below the overall average, this could indicate a more passive
behavior, as students obtain ready-made answers from the chatbot.

The average time of the Student Correction (SC) category is close to the overall
average (4.07), although the high standard deviation (8.42) indicates
significant variation. In some cases, students know the answer to an exercise
and only accomplish a quick correction from the chatbot; in others, they detect
flaws in the chatbot's responses and need to examine the subsequent answer to
identify potential errors, which suggests a more active behavior.

A message classified as Debugging Help (DH) has both a long average response
time (5.30) and high variation (10.80). Consequently, the behavior in this
category is more time-consuming and unpredictable. A possible explanation is
that some DH messages are simple, such as '\textit{what does x++ do?}'. In
contrast, others involve requests for analysis of more complex code, like
\textit{'explain this piece of code (accompanied by a complete method).'}

Conceptual Questions (CQ) generally result in a high average time (5.47) for
students to reengage with the chatbot. Moreover, the high standard deviation
(10.46) indicates an unpredictable variation in response times. This
unpredictability may be attributed to the fact that, in some cases, the
complexity of the question and answer demands more reflection time from the
student, whereas in others, lesser complexity allows for a quicker interaction
with the chatbot.

Finally, a message classified as a Complete Solution request (SCO) exhibits the
longest average time (7.65) for students to resume interaction with the chatbot.
Additionally, the high standard deviation (11.03) highlights significant
unpredictability in response time. After requesting a complete solution,
students may take longer to return due to the complexity of the problems
involved, such as the implementation of sorting algorithms such as Heapsort.

\begin{table*}[htbp]
  \caption{Average and Standard Deviation of Different Categories (in minutes)}
  \begin{center}
    \renewcommand{\arraystretch}{1.1} % Adjust the spacing between rows
    \begin{tabular}{p{4cm} p{3cm} p{3cm}} % Remove vertical bars
      \hline
      \textbf{Category} & \textbf{Average} & \textbf{Standard Deviation} \\
      \hline
      Code Snippet & 1.34 & 1.95 \\
      Uncategorized & 1.70 & 3.59 \\
      Language Change & 2.23 & 3.71 \\
      Multiple Question & 2.47 & 5.66 \\
      \hdashline
      \textit{Overall} & \textit{3.78} & \textit{6.95} \\
      \hdashline
      Student Correction & 4.07 & 8.42 \\
      Debugging Help & 5.30 & 10.80 \\
      Conceptual Questions & 5.47 & 10.46 \\
      Complete Solution & 7.65 & 11.03 \\
      \hline
    \end{tabular}
    \label{tab:averages}
  \end{center}
\end{table*}

\subsection{Interleaving}

About the third research question: \textit{RQ3 - Do students' interactions with
the chatbot alternate between study topics to suggest interleaving?}

Switching between topics can help students identify distinct concepts inside
problems. The results show that approximately 50.8\% of the conversations
include interleaving; consequently, 49.2\% are related to a single study topic.
To gain a deeper comprehension of this interleaving, conversations were analyzed
 by separating those without spacing from those that included spacing.

Conversations without spacing comprise around 66\% of the total, with 34.2\%
showing a topic change, indicating interleaving (equivalent to 22.8\% of all
conversations). Qualitative analysis of these conversations without spacing
revealed that about 7\% of the conversations presented interleaving caused by
students' at least one request to solve quizzes. However, the speed with which
students interact with the chatbot to solve quizzes may suggest a more
superficial reflection on the answers received (see section 4.2.1). On the other
hand, 8.8\% of the conversations showed no spacing but included topic changes,
suggesting that students effectively might employ a metacognitive interleaving
strategy to distinguish concepts around a problem. Finally, 7\% of the
conversations featured interaction pauses lasting between 20 and 59 minutes but
fell short of the study's definition of spacing, which required pauses of over
60 minutes. This subset of conversations revealed a natural interleaving of
topics, reflecting a chatbot usage pattern.

Conversations with spacing comprise approximately 34\% of the total, with 84.2\%
indicating interleaving (equivalent to 28\% of all conversations). Pauses longer
than 60 minutes revealed similar behavior to those shorter than 60 minutes, with
students using the chatbot primarily for quick consultations, asking questions,
and returning at least one hour late with a query about a different study topic.
The argument that some students use the chatbot primarily as a tool for quick
queries is supported by data showing an average of 4.18 messages per study
session. Interestingly, only 6\% of the conversations involved students pausing
their interaction with the chatbot for more than 60 minutes before sending a
message related to the exact topic of study.

Therefore, although approximately 50.8\% of the conversations contain
interleaving, the data shows that this topic switching is limited in terms of
using a conscious metacognitive strategy.

\section{Conclusion and Future Work}

This study aimed to analyze the interactions of Java programming students with
an educational chatbot developed using the RAG technique, employing two
metacognitive study strategies: Spacing and Interleaving.

The analysis categories, spacing, and interleaving markers provided insights
into students' interactions with the chatbot. With the concept of spacing, it
was possible to define students' study sessions, which, in this work, consisted
of a few interaction messages with the chatbot. Consequently, a significant
portion of the observed interleaving occurred due to spacing, that is, in the
transition between study sessions. These findings indicate that the intentional
use of learning strategies, such as spacing and interleaving, is still limited.
Therefore, encouraging the deliberate application of these practices can enhance
students' learning outcomes using generative artificial intelligence (GenAI)
tools.

Messages from the Code Snippets and Multiple Question (Quiz) categories stood
out for involving consistently quick interactions with the bot. This pattern
suggests some level of engagement with the chatbot but also raises the
possibility of less in-depth reflection on the content received, especially when
students request small code snippets or the resolution of exercises.

It is fundamental to consider how GenAI tools impact students in various
demographic groups, academic disciplines, cultural backgrounds, and levels of
previous experience \citep{catalan21} \citep{neo22}. Consequently, this study is
limited to a specific group of students and focuses on using a single GenAI
tool. As a result, the findings may not be generalizable to other populations or
tools.

There are ways to deepen this study; based on previous references, 60 minutes
was the standard spacing measure. However, further studies will deepen this
analysis using a quantitative variable to better represent different spacing
intervals. Future research can also include additional metacognitive strategies,
which will involve interpreting user messages in depth. For instance, the
Monitoring strategy was evident when students expressed that a concept remained
unclear and requested further clarification. Another necessary analysis involves
investigating the relationship between student profiles, metacognitive
strategies for using GenAI, and learning outcomes.

\bibliographystyle{apalike}
{\small
\bibliography{References}}

\end{document}